## QC

{r}
library(Seurat)

library(tidyverse)
library(gprofiler2)
library(DoubletFinder)
library(cowplot)



{r}
DATA <- readRDS("../Data/Seurat_Object/Mu93FB_1.rds")

{r}
DATA$SampleID <- "Mu93_1"
DATA$Condition <- "Mu93"

Start QC

{r}
DATA <- PercentageFeatureSet(DATA, "^mt-", col.name = "percent_mito")
DATA <- PercentageFeatureSet(DATA, "^Rp[sl]", col.name = "percent_ribo")
DATA <- PercentageFeatureSet(DATA, "^Hb[^(p|e|s)]", col.name = "percent_hb") #Percentage hemoglobin genes - includes all genes starting with HB except HBP.
DATA <- PercentageFeatureSet(DATA, "Pecam1|Pf4", col.name = "percent_plat")

DATA <- CellCycleScoring(DATA %>% NormalizeData(), g2m.features = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name, 
    s.features= gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name,set.ident = T)

VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "Condition", pt.size = 0.1,ncol = 3) + NoLegend()


save QC

{r save QC}
pdf("../Result/qc/QC_raw_Mu93_1.pdf", width = 10, height = 8)
VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "Condition", pt.size = 0.1,ncol = 4, alpha = 0.4) + NoLegend()
dev.off()

Check top20 genes

{r Check top20 genes}

C <- DATA@assays$RNA@layers$counts 
rownames(C) <- rownames(DATA)
colnames(C) <- colnames(DATA)
C@x <- C@x / rep.int(colSums(C), diff(C@p)) * 100

boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
#if you couldn't see the full gene names, just show the picture in a new window


From QC plots, there are some blood cells. So I would guess after I remove thoses cells, the dominant Hbb-bs here would disappear. I will check it later.

Save top20 genes

{r Save top20 genes}
pdf("../Result/qc/top20_Mu93_1.pdf", width = 12, height = 6)
boxplot(as.matrix(t(C[most_expressed, ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()
rm(C)

QC

{r QC}
dim(DATA)
#Depend on the QC plot, change the QC matrix
DATA <- subset(DATA, subset = nFeature_RNA > 200 & percent_mito < 10 & percent_hb < 10 & percent_ribo < 20 ) #,nFeature_RNA<4000 Run doubletfinder instead & percent_ribo >5 
dim(DATA) #Check cell number

[1] 28804 21221 [1] 28804 21165

DoubletFinder

Since we got much more cells than 10000, I will increase the doublet rate to 0.12.

{r doubletfinder}


DATA <- NormalizeData(DATA)
DATA <- FindVariableFeatures(DATA)
DATA <- ScaleData(DATA, vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score", "G2M.Score"))
DATA <- RunPCA(DATA, npcs = 30)
DATA <- RunUMAP(DATA, dims = 1:30)

nExp <- round(ncol(DATA) * 0.12)  # expect 8% doublets for 10000 loaded cells (Check 10X loaded chart)
DATA <- doubletFinder(DATA, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:30)
# name of the DF prediction can change, so extract the correct column name.
colnames(DATA@meta.data)[grepl("DF.classification", colnames(DATA@meta.data))]
DATA$DF <-  DATA$DF.classifications_0.25_0.09_2540
plot_grid(DimPlot(DATA, group.by = "orig.ident") + NoAxes(), 
    DimPlot(DATA, group.by = "DF") + NoAxes(), VlnPlot(DATA, features = "nFeature_RNA", group.by = "DF")
    #,FeaturePlot(DATA, features = c("Mki67","Pcna"))
    ,ncol = 2)


{r}
saveRDS(DATA, "../Data/Seurat_Object/Mu93_1_QCed.rds")

#_Mu93FB2

{r}
DATA <- readRDS("../Data/Seurat_Object/Mu93FB_2.rds")

{r}
DATA$SampleID <- "Mu93_2"
DATA$Condition <- "Mu93"

Start QC

{r}
DATA <- PercentageFeatureSet(DATA, "^mt-", col.name = "percent_mito")
DATA <- PercentageFeatureSet(DATA, "^Rp[sl]", col.name = "percent_ribo")
DATA <- PercentageFeatureSet(DATA, "^Hb[^(p|e|s)]", col.name = "percent_hb") #Percentage hemoglobin genes - includes all genes starting with HB except HBP.
DATA <- PercentageFeatureSet(DATA, "Pecam1|Pf4", col.name = "percent_plat")

DATA <- CellCycleScoring(DATA %>% NormalizeData(), g2m.features = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name, 
    s.features= gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name,set.ident = T)

VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "Condition", pt.size = 0.1,ncol = 3) + NoLegend()


save QC

{r save QC}
pdf("../Result/qc/QC_raw_Mu93_2.pdf", width = 10, height = 8)
VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "SampleID", pt.size = 0.1,ncol = 4, alpha = 0.4) + NoLegend()
dev.off()

Check top20 genes

{r Check top20 genes}

C <- DATA@assays$RNA@layers$counts 
rownames(C) <- rownames(DATA)
colnames(C) <- colnames(DATA)
C@x <- C@x / rep.int(colSums(C), diff(C@p)) * 100

boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
#if you couldn't see the full gene names, just show the picture in a new window


From QC plots, there are some blood cells. So I would guess after I remove thoses cells, the dominant Hbb-bs here would disappear. I will check it later.

Save top20 genes

{r Save top20 genes}
pdf("../Result/qc/top20_Mu93_2.pdf", width = 12, height = 6)
boxplot(as.matrix(t(C[most_expressed, ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()
rm(C)

QC

{r QC}
dim(DATA)
#Depend on the QC plot, change the QC matrix
DATA <- subset(DATA, subset = nFeature_RNA > 200 & nFeature_RNA < 7500 & nCount_RNA <40000& percent_mito < 10 & percent_hb < 10 & percent_ribo < 20 ) #,nFeature_RNA<4000 Run doubletfinder instead & percent_ribo >5 
dim(DATA) #Check cell number

[1] 29085 27145 [1] 29085 27084 [1] 29085 26929

DoubletFinder

Since we got much more cells than 10000, I will increase the doublet rate to 0.12.

{r doubletfinder}


DATA <- NormalizeData(DATA)
DATA <- FindVariableFeatures(DATA)
DATA <- ScaleData(DATA, vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score", "G2M.Score"))
DATA <- RunPCA(DATA, npcs = 30)
DATA <- RunUMAP(DATA, dims = 1:30)

nExp <- round(ncol(DATA) * 0.08)  # expect 8% doublets for 10000 loaded cells (Check 10X loaded chart), 8% for 20k cells if it is v4 chip or v3,1 HT chip.
DATA <- doubletFinder(DATA, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:30)
# name of the DF prediction can change, so extract the correct column name.
colnames(DATA@meta.data)[grepl("DF.classification", colnames(DATA@meta.data))]
DATA$DF <-  DATA$DF.classifications_0.25_0.09_2154
plot_grid(DimPlot(DATA, group.by = "orig.ident") + NoAxes(), 
    DimPlot(DATA, group.by = "DF") + NoAxes(), VlnPlot(DATA, features = "nFeature_RNA", group.by = "DF")
    #,FeaturePlot(DATA, features = c("Mki67","Pcna"))
    ,ncol = 2)


3250 to 2154

{r}
saveRDS(DATA, "../Data/Seurat_Object/Mu93_2_QCed.rds")

Mu93FB3

{r}
DATA <- readRDS("../Data/Seurat_Object/Mu93FB_3.rds")

{r}
DATA$SampleID <- "Mu93_3"
DATA$Condition <- "Mu93"

Start QC

{r}
DATA <- PercentageFeatureSet(DATA, "^mt-", col.name = "percent_mito")
DATA <- PercentageFeatureSet(DATA, "^Rp[sl]", col.name = "percent_ribo")
DATA <- PercentageFeatureSet(DATA, "^Hb[^(p|e|s)]", col.name = "percent_hb") #Percentage hemoglobin genes - includes all genes starting with HB except HBP.
DATA <- PercentageFeatureSet(DATA, "Pecam1|Pf4", col.name = "percent_plat")

DATA <- CellCycleScoring(DATA %>% NormalizeData(), g2m.features = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name, 
    s.features= gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name,set.ident = T)

VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "SampleID", pt.size = 0.1,ncol = 3) + NoLegend()


save QC

{r save QC}
pdf("../Result/qc/QC_raw_Mu93_3.pdf", width = 10, height = 8)
VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "SampleID", pt.size = 0.1,ncol = 4, alpha = 0.4) + NoLegend()
dev.off()

Check top20 genes

{r Check top20 genes}

C <- DATA@assays$RNA@layers$counts 
rownames(C) <- rownames(DATA)
colnames(C) <- colnames(DATA)
C@x <- C@x / rep.int(colSums(C), diff(C@p)) * 100

boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
#if you couldn't see the full gene names, just show the picture in a new window


From QC plots, there are some blood cells. So I would guess after I remove thoses cells, the dominant Hbb-bs here would disappear. I will check it later.

Save top20 genes

{r Save top20 genes}
pdf("../Result/qc/top20_Mu93_3.pdf", width = 12, height = 6)
boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()
rm(C)

QC

{r QC}
dim(DATA)
#Depend on the QC plot, change the QC matrix
DATA <- subset(DATA, subset = nFeature_RNA > 200 & nFeature_RNA < 7500 & nCount_RNA <40000 & percent_mito < 10 & percent_hb < 10 & percent_ribo < 20 ) #,nFeature_RNA<4000 Run doubletfinder instead & percent_ribo >5 
dim(DATA) #Check cell number

[1] 28724 26465 [1] 28724 26447 [1] 28724 26345

DoubletFinder

Since we got much more cells than 10000, I will increase the doublet rate to 0.12.

{r doubletfinder}

DATA <- NormalizeData(DATA)
DATA <- FindVariableFeatures(DATA)
DATA <- ScaleData(DATA, vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score", "G2M.Score"))
DATA <- RunPCA(DATA, npcs = 30)
DATA <- RunUMAP(DATA, dims = 1:30)

nExp <- round(ncol(DATA) * 0.08)  # expect 8% doublets for 10000 loaded cells (Check 10X loaded chart)
DATA <- doubletFinder(DATA, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:30)
# name of the DF prediction can change, so extract the correct column name.
colnames(DATA@meta.data)[grepl("DF.classification", colnames(DATA@meta.data))]
DATA$DF <-  DATA$DF.classifications_0.25_0.09_2108
plot_grid(DimPlot(DATA, group.by = "orig.ident") + NoAxes(), 
    DimPlot(DATA, group.by = "DF") + NoAxes(), VlnPlot(DATA, features = "nFeature_RNA", group.by = "DF")
    #,FeaturePlot(DATA, features = c("Mki67","Pcna"))
    ,ncol = 2)


3174 to 2108

{r}
saveRDS(DATA, "../Data/Seurat_Object/Mu93_3_QCed.rds")

MuLI1

{r}
DATA <- readRDS("../Data/Seurat_Object/MuLIFB_1.rds")

{r}
DATA$SampleID <- "MuLI_1"
DATA$Condition <- "MuLI"

Start QC

{r}
DATA <- PercentageFeatureSet(DATA, "^mt-", col.name = "percent_mito")
DATA <- PercentageFeatureSet(DATA, "^Rp[sl]", col.name = "percent_ribo")
DATA <- PercentageFeatureSet(DATA, "^Hb[^(p|e|s)]", col.name = "percent_hb") #Percentage hemoglobin genes - includes all genes starting with HB except HBP.
DATA <- PercentageFeatureSet(DATA, "Pecam1|Pf4", col.name = "percent_plat")

DATA <- CellCycleScoring(DATA %>% NormalizeData(), g2m.features = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name, 
    s.features= gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name,set.ident = T)

VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "Condition", pt.size = 0.1,ncol = 3) + NoLegend()


save QC

{r save QC}
pdf("../Result/qc/QC_raw_MuLI_1.pdf", width = 10, height = 8)
VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "Condition", pt.size = 0.1,ncol = 4, alpha = 0.4) + NoLegend()
dev.off()

Check top20 genes

{r Check top20 genes}

C <- DATA@assays$RNA@layers$counts 
rownames(C) <- rownames(DATA)
colnames(C) <- colnames(DATA)
C@x <- C@x / rep.int(colSums(C), diff(C@p)) * 100

boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
#if you couldn't see the full gene names, just show the picture in a new window


From QC plots, there are some blood cells. So I would guess after I remove thoses cells, the dominant Hbb-bs here would disappear. I will check it later.

Save top20 genes

{r Save top20 genes}
pdf("../Result/qc/top20_MuLI_1.pdf", width = 12, height = 6)
boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()
rm(C)

QC

{r QC}
dim(DATA)
#Depend on the QC plot, change the QC matrix
DATA <- subset(DATA, subset = nFeature_RNA > 200  & nFeature_RNA < 7500 & nCount_RNA <40000&  percent_mito < 10 & percent_hb < 10 & percent_ribo < 20 ) #,nFeature_RNA<4000 Run doubletfinder instead & percent_ribo >5 
dim(DATA) #Check cell number

[1] 28426 19409 [1] 28426 19378 [1] 28426 19324

DoubletFinder

Since we got much more cells than 10000, I will increase the doublet rate to 0.12.

{r doubletfinder}


DATA <- NormalizeData(DATA)
DATA <- FindVariableFeatures(DATA)
DATA <- ScaleData(DATA, vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score", "G2M.Score"))
DATA <- RunPCA(DATA, npcs = 30)
DATA <- RunUMAP(DATA, dims = 1:30)

nExp <- round(ncol(DATA) * 0.08)  # expect 8% doublets for 10000 loaded cells (Check 10X loaded chart)
DATA <- doubletFinder(DATA, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:30)
# name of the DF prediction can change, so extract the correct column name.
colnames(DATA@meta.data)[grepl("DF.classification", colnames(DATA@meta.data))]
DATA$DF <-  DATA$DF.classifications_0.25_0.09_1546
plot_grid(DimPlot(DATA, group.by = "orig.ident") + NoAxes(), 
    DimPlot(DATA, group.by = "DF") + NoAxes(), VlnPlot(DATA, features = "nFeature_RNA", group.by = "DF")
    #,FeaturePlot(DATA, features = c("Mki67","Pcna"))
    ,ncol = 2)


2325 to 1546

{r}
saveRDS(DATA, "../Data/Seurat_Object/MuLI_1_QCed.rds")

MuLIFB2

{r}
DATA <- readRDS("../Data/Seurat_Object/MuLIFB_2.rds")

{r}
DATA$SampleID <- "MuLI_2"
DATA$Condition <- "MuLI"

Start QC

{r}
DATA <- PercentageFeatureSet(DATA, "^mt-", col.name = "percent_mito")
DATA <- PercentageFeatureSet(DATA, "^Rp[sl]", col.name = "percent_ribo")
DATA <- PercentageFeatureSet(DATA, "^Hb[^(p|e|s)]", col.name = "percent_hb") #Percentage hemoglobin genes - includes all genes starting with HB except HBP.
DATA <- PercentageFeatureSet(DATA, "Pecam1|Pf4", col.name = "percent_plat")

DATA <- CellCycleScoring(DATA %>% NormalizeData(), g2m.features = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name, 
    s.features= gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name,set.ident = T)

VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "SampleID", pt.size = 0.1,ncol = 3) + NoLegend()


save QC

{r save QC}
pdf("../Result/qc/QC_raw_MuLI_2.pdf", width = 10, height = 8)
VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "SampleID", pt.size = 0.1,ncol = 4, alpha = 0.4) + NoLegend()
dev.off()

Check top20 genes

{r Check top20 genes}

C <- DATA@assays$RNA@layers$counts 
rownames(C) <- rownames(DATA)
colnames(C) <- colnames(DATA)
C@x <- C@x / rep.int(colSums(C), diff(C@p)) * 100

boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
#if you couldn't see the full gene names, just show the picture in a new window


From QC plots, there are some blood cells. So I would guess after I remove thoses cells, the dominant Hbb-bs here would disappear. I will check it later.

Save top20 genes

{r Save top20 genes}
pdf("../Result/qc/top20_MuLI_2.pdf", width = 12, height = 6)
boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()
rm(C)

QC

{r QC}
dim(DATA)
#Depend on the QC plot, change the QC matrix
DATA <- subset(DATA, subset = nFeature_RNA > 200 & percent_mito < 10 & percent_hb < 10 & percent_ribo < 20 ) #,nFeature_RNA<4000 Run doubletfinder instead & percent_ribo >5 
dim(DATA) #Check cell number

[1] 28629 22307 [1] 28629 22286

DoubletFinder

Since we got much more cells than 10000, I will increase the doublet rate to 0.12.

{r doubletfinder}


DATA <- NormalizeData(DATA)
DATA <- FindVariableFeatures(DATA)
DATA <- ScaleData(DATA, vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score", "G2M.Score"))
DATA <- RunPCA(DATA, npcs = 30)
DATA <- RunUMAP(DATA, dims = 1:30)

nExp <- round(ncol(DATA) * 0.12)  # expect 8% doublets for 10000 loaded cells (Check 10X loaded chart)
DATA <- doubletFinder(DATA, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:30)
# name of the DF prediction can change, so extract the correct column name.
colnames(DATA@meta.data)[grepl("DF.classification", colnames(DATA@meta.data))]
DATA$DF <-  DATA$DF.classifications_0.25_0.09_2674
plot_grid(DimPlot(DATA, group.by = "orig.ident") + NoAxes(), 
    DimPlot(DATA, group.by = "DF") + NoAxes(), VlnPlot(DATA, features = "nFeature_RNA", group.by = "DF")
    #,FeaturePlot(DATA, features = c("Mki67","Pcna"))
    ,ncol = 2)


{r}
saveRDS(DATA, "../Data/Seurat_Object/MuLI_2_QCed.rds")

MuLIFB3

{r}
DATA <- readRDS("../Data/Seurat_Object/MuLIFB_3.rds")

{r}
DATA$SampleID <- "MuLI_3"
DATA$Condition <- "MuLI"

Start QC

{r}
DATA <- PercentageFeatureSet(DATA, "^mt-", col.name = "percent_mito")
DATA <- PercentageFeatureSet(DATA, "^Rp[sl]", col.name = "percent_ribo")
DATA <- PercentageFeatureSet(DATA, "^Hb[^(p|e|s)]", col.name = "percent_hb") #Percentage hemoglobin genes - includes all genes starting with HB except HBP.
DATA <- PercentageFeatureSet(DATA, "Pecam1|Pf4", col.name = "percent_plat")

DATA <- CellCycleScoring(DATA %>% NormalizeData(), g2m.features = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name, 
    s.features= gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name,set.ident = T)

VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "Condition", pt.size = 0.1,ncol = 3) + NoLegend()


save QC

{r save QC}
pdf("../Result/qc/QC_raw_MuLI_3.pdf", width = 10, height = 8)
VlnPlot(DATA, features = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb", "percent_plat","S.Score", "G2M.Score"),group.by = "SampleID", pt.size = 0.1,ncol = 4, alpha = 0.4) + NoLegend()
dev.off()

Check top20 genes

{r Check top20 genes}

C <- DATA@assays$RNA@layers$counts 
rownames(C) <- rownames(DATA)
colnames(C) <- colnames(DATA)
C@x <- C@x / rep.int(colSums(C), diff(C@p)) * 100

boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
#if you couldn't see the full gene names, just show the picture in a new window


From QC plots, there are some blood cells. So I would guess after I remove thoses cells, the dominant Hbb-bs here would disappear. I will check it later.

Save top20 genes

{r Save top20 genes}
pdf("../Result/qc/top20_MuLI_3.pdf", width = 12, height = 6)
boxplot(as.matrix(t(C[order(Matrix::rowSums(C), decreasing = T)[20:1], ])),
    cex = 0.1, las = 1, xlab = "Percent counts per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()
rm(C)

QC

{r QC}
dim(DATA)
#Depend on the QC plot, change the QC matrix
DATA <- subset(DATA, subset = nFeature_RNA > 200   & nFeature_RNA < 7500 & nCount_RNA <40000&   percent_mito < 10 & percent_hb < 10 & percent_ribo < 20 ) #,nFeature_RNA<4000 Run doubletfinder instead & percent_ribo >5 
dim(DATA) #Check cell number

[1] 28639 24321 [1] 28639 24309 [1] 28639 24263

DoubletFinder

Since we got much more cells than 10000, I will increase the doublet rate to 0.12.

{r doubletfinder}


DATA <- NormalizeData(DATA)
DATA <- FindVariableFeatures(DATA)
DATA <- ScaleData(DATA, vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score", "G2M.Score"))
DATA <- RunPCA(DATA, npcs = 30)
DATA <- RunUMAP(DATA, dims = 1:30)

nExp <- round(ncol(DATA) * 0.08)  # expect 8% doublets for 10000 loaded cells (Check 10X loaded chart)
DATA <- doubletFinder(DATA, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:30)
# name of the DF prediction can change, so extract the correct column name.
colnames(DATA@meta.data)[grepl("DF.classification", colnames(DATA@meta.data))]
DATA$DF <-  DATA$DF.classifications_0.25_0.09_1941
plot_grid(DimPlot(DATA, group.by = "orig.ident") + NoAxes(), 
    DimPlot(DATA, group.by = "DF") + NoAxes(), VlnPlot(DATA, features = "nFeature_RNA", group.by = "DF")
    #,FeaturePlot(DATA, features = c("Mki67","Pcna"))
    ,ncol = 2)


2917 to 1941

{r}
saveRDS(DATA, "../Data/Seurat_Object/MuLI_3_QCed.rds")


## Integration
DATA[['RNA']] <- split(DATA[['RNA']], f = DATA$SampleID)

DATA <- NormalizeData(DATA)
DATA <- FindVariableFeatures(DATA)
VariableFeatures(DATA) = SelectIntegrationFeatures5(DATA, nfeatures = 3000)
DATA <- ScaleData(DATA,vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score","G2M.Score"))

#DATA <- SCTransform(DATA,vars.to.regress = c("nFeature_RNA","nCount_RNA","percent_mito","percent_ribo", "percent_hb","percent_plat","S.Score","G2M.Score") )


DATA <- RunPCA(DATA,npcs = 50)

#unintegrated
#DATA <- FindNeighbors(DATA, dims = 1:30, reduction = "pca")
DATA <- RunUMAP(DATA, dims = 1:50, reduction = "pca", reduction.name = "umap.unintegrated")
p1<-DimPlot(DATA, reduction = "umap.unintegrated",shuffle = T,label = T)+ggtitle("unintegrated")+NoLegend()
#CCA
DATA <- IntegrateLayers(
  object = DATA, method = CCAIntegration,
  orig.reduction = "pca", new.reduction = "integrated.cca")
DATA <- FindNeighbors(DATA, reduction = "integrated.cca", dims = 1:30)
DATA <- RunUMAP(DATA, reduction = "integrated.cca", dims = 1:50, reduction.name = "umap.cca")

#Harmony
DATA <- IntegrateLayers(
  object = DATA, method = HarmonyIntegration,
  orig.reduction = "pca", new.reduction = "harmony"
)
#DATA <- FindNeighbors(DATA, reduction = "harmony", dims = 1:30)
DATA <- RunUMAP(DATA, reduction = "harmony", dims = 1:40, reduction.name = "umap.harmony")
DimPlot(DATA,group.by = "Subtypes2", reduction = "umap.harmony", shuffle = T,label = T)+ggtitle("Harmony")

## hdWGCNA

```{bash}
ml PDC/24.11
ml R/4.4.2 Rbio/3.20-cpeGNU-24.11

```

```{r}
.libPaths("/cfs/klemming/projects/supr/naiss2023-6-180/Yuxi/R_lib")
#install.packages("ggplot2")
#install.packages("rlang")
library(Seurat)
library(tidyverse)
library(WGCNA)
library(hdWGCNA)
```

You can add options to executable code like this

```{r}
DATA <- readRDS("DATA_final.rds")
DATA <- SetIdent(DATA, value = "Celltypes")
DATA$Celltypes <-factor(x= DATA$Celltypes, levels =  c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs","MG"))
DATA$Celltypes[WhichCells(DATA, idents = "MG")] <- "Microglia"
DATA@reductions$UMAP <- DATA@reductions$umap.harmony30
DATA@reductions$umap.harmony30 <- NULL
DATA$Celltypes <-factor(x= DATA$Celltypes, levels =  c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs"))
```

## IbNs

```{r}
#DATA <- readRDS("wgcna.rds") %>% JoinLayers()
enableWGCNAThreads(nThreads = 8)
DATA <- SetupForWGCNA(
  DATA,
  gene_select = "fraction", # the gene selection approach
  fraction = 0.05, # fraction of cells that a gene needs to be expressed in order to be included
  wgcna_name = "AD_treatment" # the name of the hdWGCNA experiment
)

DATA <- MetacellsByGroups(DATA, group.by=c("Celltypes", "SampleID"), reduction = "harmony",k = 25, max_shared = 10, ident.group = "Celltypes")

table(DATA$metacell_grouping)
DATA <- NormalizeMetacells(DATA)
DATA <- SetDatExpr(DATA,  group_name= "IbNs",group.by = "Celltypes",assay = "RNA", layer = 'data')
DATA <- TestSoftPowers(DATA, networkType = 'signed')
plot_list <- PlotSoftPowers(DATA)

patchwork::wrap_plots(plot_list, ncol=2)
#saveRDS(DATA, "wgcna.rds")
```

```{r}
DATA <- readRDS("wgcna.rds")

power_table <- GetPowerTable(DATA)
power_table
```

```{r}
DATA <- ConstructNetwork(
  DATA,
  tom_name = 'IbNs_tom', # name of the topoligical overlap matrix written to disk
  overwrite_tom = T
)

pdf("IbNs Dedrogram.pdf", width = 8, height = 4)
PlotDendrogram(DATA, main='IbNs hdWGCNA Nodule Dendrogram')
dev.off()

```

Soft_power = 5.

```{r}
# need to run ScaleData first or else harmony throws an error:
DATA <- ScaleData(DATAj, features=VariableFeatures(DATA))

# compute all MEs in the full single-cell dataset
DATA <- ModuleEigengenes(
 DATA,
 group.by.vars="SampleID"
)
```

```{r}


DATA <- ModuleConnectivity(DATA, group.by = "Celltypes",group_name = 'IbNs')
DATA <- ResetModuleNames(
  DATA,
  new_name = "IbNs-M"
)
PlotKMEs(DATA, ncol=5)

library(UCell)
DATA <- ModuleExprScore(
  DATA,
  n_genes = 25,
  method='UCell'
)

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='hMEs', # plot the hMEs
  order=TRUE # order so the points with highest hMEs are on top
)

# stitch together with patchwork
pdf("hMEs_IbNs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='scores', # plot the hub gene scores
  order='shuffle', # order so cells are shuffled
  ucell = TRUE # depending on Seurat vs UCell for gene scoring
)
pdf("Hub_IbNs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

# stitch together with patchwork
wrap_plots(plot_list, ncol=6)

saveRDS(DATA, "IbNs_wgcna.rds")
```

```{r}
modules <- GetModules(DATA)
mods <- levels(modules$module); mods <- mods[mods != 'grey']

# hubgene network
g<-HubGeneNetworkPlot(
  DATA,
  n_hubs = 10, n_other=30,
  edge_prop = 0.75,
  mods = mods[c(9)], return_graph = T) 

```

```{r}

pdf("IbN-M5_net.pdf", height = 4, width = 4)
# 2. Plot using ggraph (which uses standard ggplot2 syntax)
ggraph(g, layout = "fr") + # 'fr' is a standard force-directed layout
  
  # --- 1. CHANGE EDGE COLOR ---
  geom_edge_link(
    color = "#67a8cd",     # Set your custom edge color here
    alpha = 0.4,            # Adjust transparency
    width = 0.5
  ) + 
  
  # --- 2. CHANGE NODE COLOR ---
  geom_node_point(
    color = "#67a8cd",    # Set your custom node color here
    size = 2.5               # Adjust node size
  ) + 
  
  # --- 3. ADD A BOX FOR THE LABELS ---
  geom_node_label(
    aes(label = label),      # 'name' is usually the default igraph attribute for gene names
    fill = "white",         # Background color of the label box
    color = "black",        # Text color inside the box
    repel = T,           # Pushes labels apart so they don't overlap
    label.padding = unit(0.2, "lines"), # Size of the box around the text
    label.r = unit(0.2, "lines"),       # Rounded corners for the box
    size = 2
  ) +
  
  # Clean up the background
  theme_graph()
dev.off()
```

```{r}
options(repr.plot.width=4, repr.plot.height=39)


plots <- list()


interest = c("IbNs-M3", "IbNs-M5", "IbNs-M9")
# Replace this with your actual process of creating each graph
for (i in 1:length(interest)) {
  # Assuming `cur_network` is defined and updated for each graph creation
  graph <- subset(cur_network, module1 == interest[i]) %>% 
    igraph::graph_from_data_frame() %>%
    tidygraph::as_tbl_graph(directed = FALSE) %>% 
    tidygraph::activate(nodes)
  
  # Add the module name to the graph
  V(graph)$module <- interest[i]
  
  # Get the top 25 hub genes for each module
  hub_genes <- GetHubGenes(DATA, n_hubs = 25) %>% .
hub <- ifelse(V(graph)$name %in% hub_genes, V(graph)$name, "")
  
  # Make the plot with ggraph
  p <- ggraph(graph) + 
    geom_edge_link(aes(alpha = weight, color = edge_color)) + 
    geom_node_point(aes(color = module, size = 3)) +
    geom_node_label(aes(label = hub), repel = TRUE, max.overlaps = Inf, fontface = 'italic', size = 6) +
    scale_colour_manual(values = mod_cp) +
    scale_edge_colour_manual(values = mod_cp) +
    theme_void() + # use theme_void() instead of theme_nothing() to avoid errors
    theme(legend.position = "none") # use theme(legend.position = "none") instead of NoLegend()
  
  # Store the plot in the list
  plots[[i]] <- p
}

plots <- lapply(plots, function(p) p + theme(plot.margin = margin(5, 5, 5, 5))) 
grid.arrange(grobs = plots, nrow = 11, ncol = 1)
```

```{r}

DATA <- readRDS("IbNs_wgcna.rds")

DATA <- RunModuleUMAP(
  DATA,
  n_hubs = 10, # number of hub genes to include for the UMAP embedding
  n_neighbors=15, # neighbors parameter for UMAP
  min_dist=0.1 # min distance between points in UMAP space
)

pdf("IbNs_ModuleUMAP.pdf", width = 10, height = 6)
ModuleUMAPPlot(DATA,
  edge.alpha=0.25,
  sample_edges=TRUE,
  edge_prop=0.1, # proportion of edges to sample (20% here)
  label_hubs=2 ,# how many hub genes to plot per module?
  keep_grey_edges=FALSE)
dev.off()

Ctrl <- DATA@meta.data %>% subset(Celltypes == 'IbNs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'IbNs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

head(DMEs)

pdf("DME_lolli_IbNs.pdf", width =10, height = 7)
PlotDMEsLollipop(
  DATA, 
  DMEs, 
  wgcna_name='AD_treatment', 
  pvalue = "p_val_adj")
dev.off()

pdf("DME_vol_IbNs.pdf", width =10, height = 7)
PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)
dev.off()
```

### Enrichment

```{r}
library(enrichR)
library(GeneOverlap)
```

```{r}
DATA <- RunEnrichr(
  DATA,
  dbs=c("GO_Biological_Process_2025", "Reactome_Pathways_2024"),
  max_genes = 100 # use max_genes = Inf to choose all genes
)

# retrieve the output table
enrich_df <- GetEnrichrTable(DATA)

# look at the results
head(enrich_df)
write.csv(enrich_df, "IbNs_enrich.csv")
pdf("IbNs_enrich.pdf")
EnrichrBarPlot(
  DATA,
  outdir = "enrichr_plots", # name of output directory
  n_terms = 10, # number of enriched terms to show (sometimes more are shown if there are ties)
  plot_size = c(5,7), # width, height of the output .pdfs
  logscale=TRUE # do you want to show the enrichment as a log scale?
)
dev.off()
```

## ExNs

```{r}
DATA <- readRDS("DATA_final.rds") 
DATA <- SetIdent(DATA, value = "Celltypes") 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs","MG")) 
DATA$Celltypes[WhichCells(DATA, idents = "MG")] <- "Microglia"
DATA@reductions$UMAP <- DATA@reductions$umap.harmony30
DATA@reductions$umap.harmony30 <- NULL 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs"))

DATA <- SetupForWGCNA(
  DATA,
  gene_select = "fraction", # the gene selection approach
  fraction = 0.05, # fraction of cells that a gene needs to be expressed in order to be included
  wgcna_name = "AD_treatment" # the name of the hdWGCNA experiment
)

DATA <- MetacellsByGroups(DATA, group.by=c("Celltypes", "SampleID"), reduction = "harmony",k = 25, max_shared = 10, ident.group = "Celltypes")

table(DATA$metacell_grouping)
DATA <- NormalizeMetacells(DATA)
DATA <- SetDatExpr(DATA,  group_name= "ExNs",group.by = "Celltypes",assay = "RNA", layer = 'data')
DATA <- TestSoftPowers(DATA, networkType = 'signed')
plot_list <- PlotSoftPowers(DATA)

patchwork::wrap_plots(plot_list, ncol=2)

DATA <- ConstructNetwork(
  DATA,
  tom_name = 'ExNs_tom', # name of the topoligical overlap matrix written to disk
  overwrite_tom = T
) #soft power =7

# compute all MEs in the full single-cell dataset
DATA <- ModuleEigengenes(
 DATA,
 group.by.vars="SampleID"
)

DATA <- ModuleConnectivity(DATA, group.by = "Celltypes",group_name = 'ExNs')
DATA <- ResetModuleNames(
  DATA,
  new_name = "ExNs-M"
)
PlotKMEs(DATA, ncol=5)

library(UCell)
DATA <- ModuleExprScore(
  DATA,
  n_genes = 25,
  method='UCell'
)

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='hMEs', # plot the hMEs
  order=TRUE # order so the points with highest hMEs are on top
)

# stitch together with patchwork
pdf("hMEs_ExNs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='scores', # plot the hub gene scores
  order='shuffle', # order so cells are shuffled
  ucell = TRUE # depending on Seurat vs UCell for gene scoring
)
pdf("Hub_ExNs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

# stitch together with patchwork
wrap_plots(plot_list, ncol=6)

saveRDS(DATA, "ExNs_wgcna.rds")

DATA <- readRDS("ExNs_wgcna.rds")
Ctrl <- DATA@meta.data %>% subset(Celltypes == 'ExNs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'ExNs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

head(DMEs)

pdf("DME_lolli_ExNs.pdf", width =10, height = 7)
PlotDMEsLollipop(
  DATA, 
  DMEs, 
  wgcna_name='AD_treatment', 
  pvalue = "p_val_adj")
dev.off()

pdf("DME_vol_ExNs.pdf", width =10, height = 7)
PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)
dev.off()
```


```{r}



modules <- GetModules(DATA)
mods <- levels(modules$module); mods <- mods[mods != 'grey']

# hubgene network
g<-HubGeneNetworkPlot(
  DATA,
  n_hubs = 10, n_other=30,
  edge_prop = 0.75,
  mods = mods[c(15)], return_graph = T) 

pdf("ExN-M5_net.pdf", height = 4, width = 4)
# 2. Plot using ggraph (which uses standard ggplot2 syntax)
ggraph(g, layout = "fr") + # 'fr' is a standard force-directed layout
  
  # --- 1. CHANGE EDGE COLOR ---
  geom_edge_link(
    color = "#f36569",     # Set your custom edge color here
    alpha = 0.4,            # Adjust transparency
    width = 0.5
  ) + 
  
  # --- 2. CHANGE NODE COLOR ---
  geom_node_point(
    color = "#f36569",    # Set your custom node color here
    size = 2.5               # Adjust node size
  ) + 
  
  # --- 3. ADD A BOX FOR THE LABELS ---
  geom_node_label(
    aes(label = label),      # 'name' is usually the default igraph attribute for gene names
    fill = "white",         # Background color of the label box
    color = "black",        # Text color inside the box
    repel = T,           # Pushes labels apart so they don't overlap
    label.padding = unit(0.2, "lines"), # Size of the box around the text
    label.r = unit(0.2, "lines"),       # Rounded corners for the box
    size = 1.5
  ) +
  
  # Clean up the background
  theme_graph()
dev.off()
```

### Enrichment

```{r}
DATA <- RunEnrichr(
  DATA,
  dbs=c("GO_Biological_Process_2025", "Reactome_Pathways_2024"),
  max_genes = 100 # use max_genes = Inf to choose all genes
)

# retrieve the output table
enrich_df <- GetEnrichrTable(DATA)

# look at the results
head(enrich_df)
write.csv(enrich_df, "ExNs_enrich.csv")
pdf("ExNs_enrich.pdf")
EnrichrBarPlot(
  DATA,
  outdir = "enrichr_plots", # name of output directory
  n_terms = 10, # number of enriched terms to show (sometimes more are shown if there are ties)
  plot_size = c(5,7), # width, height of the output .pdfs
  logscale=TRUE # do you want to show the enrichment as a log scale?
)
dev.off()

```

## MG

```{r}
DATA <- readRDS("DATA_final.rds") 
DATA <- SetIdent(DATA, value = "Celltypes") 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs","MG")) 
DATA$Celltypes[WhichCells(DATA, idents = "MG")] <- "Microglia"
DATA@reductions$UMAP <- DATA@reductions$umap.harmony30
DATA@reductions$umap.harmony30 <- NULL 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs"))

DATA <- SetupForWGCNA(
  DATA,
  gene_select = "fraction", # the gene selection approach
  fraction = 0.05, # fraction of cells that a gene needs to be expressed in order to be included
  wgcna_name = "AD_treatment" # the name of the hdWGCNA experiment
)

DATA <- MetacellsByGroups(DATA, group.by=c("Celltypes", "SampleID"), reduction = "harmony",k = 25, max_shared = 10, ident.group = "Celltypes")

table(DATA$metacell_grouping)
DATA <- NormalizeMetacells(DATA)
DATA <- SetDatExpr(DATA,  group_name= "Microglia",group.by = "Celltypes",assay = "RNA", layer = 'data')
DATA <- TestSoftPowers(DATA, networkType = 'signed')
plot_list <- PlotSoftPowers(DATA)

patchwork::wrap_plots(plot_list, ncol=2)

DATA <- ConstructNetwork(
  DATA,
  tom_name = 'Microglia_tom', # name of the topoligical overlap matrix written to disk
  overwrite_tom = T
) #soft power =7

# compute all MEs in the full single-cell dataset
DATA <- ModuleEigengenes(
 DATA,
 group.by.vars="SampleID"
)

DATA <- ModuleConnectivity(DATA, group.by = "Celltypes",group_name = 'Microglia')
DATA <- ResetModuleNames(
  DATA,
  new_name = "Microglia-M"
)
PlotKMEs(DATA, ncol=5)

library(UCell)
DATA <- ModuleExprScore(
  DATA,
  n_genes = 25,
  method='UCell'
)

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='hMEs', # plot the hMEs
  order=TRUE # order so the points with highest hMEs are on top
)

# stitch together with patchwork
pdf("hMEs_Microglia.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='scores', # plot the hub gene scores
  order='shuffle', # order so cells are shuffled
  ucell = TRUE # depending on Seurat vs UCell for gene scoring
)
pdf("Hub_Microglia.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

# stitch together with patchwork
wrap_plots(plot_list, ncol=6)

saveRDS(DATA, "Microglia_wgcna.rds")
DATA <- readRDS("Microglia_wgcna.rds")

Ctrl <- DATA@meta.data %>% subset(Celltypes == 'Microglia' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'Microglia' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

head(DMEs)

pdf("DME_lolli_microglia.pdf", width =10, height = 7)
PlotDMEsLollipop(
  DATA, 
  DMEs, 
  wgcna_name='AD_treatment', 
  pvalue = "p_val_adj")
dev.off()

pdf("DME_vol_microglia.pdf", width =10, height = 7)
PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)
dev.off()
```

### Enrichment

```{r}

DATA <- RunEnrichr(
  DATA,
  dbs=c("GO_Biological_Process_2025", "Reactome_Pathways_2024"),
  max_genes = 100 # use max_genes = Inf to choose all genes
)

# retrieve the output table
enrich_df <- GetEnrichrTable(DATA)

# look at the results
head(enrich_df)
write.csv(enrich_df, "Microglia_enrich.csv")
pdf("Microglia_enrich.pdf")
EnrichrBarPlot(
  DATA,
  outdir = "enrichr_plots", # name of output directory
  n_terms = 10, # number of enriched terms to show (sometimes more are shown if there are ties)
  plot_size = c(5,7), # width, height of the output .pdfs
  logscale=TRUE # do you want to show the enrichment as a log scale?
)
dev.off()

```

## ASCs

```{r}
DATA <- readRDS("DATA_final.rds") 
DATA <- SetIdent(DATA, value = "Celltypes") 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs","MG")) 
DATA$Celltypes[WhichCells(DATA, idents = "MG")] <- "Microglia"
DATA@reductions$UMAP <- DATA@reductions$umap.harmony30
DATA@reductions$umap.harmony30 <- NULL 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs"))

DATA <- SetupForWGCNA(
  DATA,
  gene_select = "fraction", # the gene selection approach
  fraction = 0.05, # fraction of cells that a gene needs to be expressed in order to be included
  wgcna_name = "AD_treatment" # the name of the hdWGCNA experiment
)

DATA <- MetacellsByGroups(DATA, group.by=c("Celltypes", "SampleID"), reduction = "harmony",k = 25, max_shared = 10, ident.group = "Celltypes")

table(DATA$metacell_grouping)
DATA <- NormalizeMetacells(DATA)
DATA <- SetDatExpr(DATA,  group_name= "ASCs",group.by = "Celltypes",assay = "RNA", layer = 'data')
DATA <- TestSoftPowers(DATA, networkType = 'signed')
plot_list <- PlotSoftPowers(DATA)

patchwork::wrap_plots(plot_list, ncol=2)

DATA <- ConstructNetwork(
  DATA,
  tom_name = 'ASCs_tom', # name of the topoligical overlap matrix written to disk
  overwrite_tom = T
) #soft power =7

# compute all MEs in the full single-cell dataset
DATA <- ModuleEigengenes(
 DATA,
 group.by.vars="SampleID"
)

DATA <- ModuleConnectivity(DATA, group.by = "Celltypes",group_name = 'ASCs')
DATA <- ResetModuleNames(
  DATA,
  new_name = "ASCs-M"
)
PlotKMEs(DATA, ncol=5)

library(UCell)
DATA <- ModuleExprScore(
  DATA,
  n_genes = 25,
  method='UCell'
)

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='hMEs', # plot the hMEs
  order=TRUE # order so the points with highest hMEs are on top
)

# stitch together with patchwork
pdf("hMEs_ASCs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='scores', # plot the hub gene scores
  order='shuffle', # order so cells are shuffled
  ucell = TRUE # depending on Seurat vs UCell for gene scoring
)
pdf("Hub_ASCs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

# stitch together with patchwork
patchwork::wrap_plots(plot_list, ncol=6)

saveRDS(DATA, "ASCs_wgcna.rds")
#DATA <- readRDS("Microglia_wgcna.rds")

Ctrl <- DATA@meta.data %>% subset(Celltypes == 'ASCs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'ASCs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

head(DMEs)

pdf("DME_lolli_ASCs.pdf", width =10, height = 7)
PlotDMEsLollipop(
  DATA, 
  DMEs, 
  wgcna_name='AD_treatment', 
  pvalue = "p_val_adj")
dev.off()

pdf("DME_vol_ASCs.pdf", width =10, height = 7)
PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)
dev.off()
```

### Enrichment

```{r}

DATA <- RunEnrichr(
  DATA,
  dbs=c("GO_Biological_Process_2025", "Reactome_Pathways_2024"),
  max_genes = Inf # use max_genes = Inf to choose all genes
)

# retrieve the output table
enrich_df <- GetEnrichrTable(DATA)

# look at the results
head(enrich_df)
write.csv(enrich_df, "ASCs_enrich.csv")
pdf("ASCs_enrich.pdf")
EnrichrBarPlot(
  DATA,
  outdir = "enrichr_plots", # name of output directory
  n_terms = 10, # number of enriched terms to show (sometimes more are shown if there are ties)
  plot_size = c(5,7), # width, height of the output .pdfs
  logscale=TRUE # do you want to show the enrichment as a log scale?
)
dev.off()

```

## OLs

```{r}
DATA <- readRDS("DATA_final.rds") 
DATA <- SetIdent(DATA, value = "Celltypes") 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs","MG")) 
DATA$Celltypes[WhichCells(DATA, idents = "MG")] <- "Microglia"
DATA@reductions$UMAP <- DATA@reductions$umap.harmony30
DATA@reductions$umap.harmony30 <- NULL 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs"))

DATA <- SetupForWGCNA(
  DATA,
  gene_select = "fraction", # the gene selection approach
  fraction = 0.05, # fraction of cells that a gene needs to be expressed in order to be included
  wgcna_name = "AD_treatment" # the name of the hdWGCNA experiment
)

DATA <- MetacellsByGroups(DATA, group.by=c("Celltypes", "SampleID"), reduction = "harmony",k = 25, max_shared = 10, ident.group = "Celltypes")

table(DATA$metacell_grouping)
DATA <- NormalizeMetacells(DATA)
DATA <- SetDatExpr(DATA,  group_name= "OLs",group.by = "Celltypes",assay = "RNA", layer = 'data')
DATA <- TestSoftPowers(DATA, networkType = 'signed')
plot_list <- PlotSoftPowers(DATA)

patchwork::wrap_plots(plot_list, ncol=2)

DATA <- ConstructNetwork(
  DATA,
  tom_name = 'OLs_tom', # name of the topoligical overlap matrix written to disk
  overwrite_tom = T
) #soft power =7

# compute all MEs in the full single-cell dataset
DATA <- ModuleEigengenes(
 DATA,
 group.by.vars="SampleID"
)

DATA <- ModuleConnectivity(DATA, group.by = "Celltypes",group_name = 'OLs')
DATA <- ResetModuleNames(
  DATA,
  new_name = "OLs-M"
)
PlotKMEs(DATA, ncol=5)

library(UCell)
DATA <- ModuleExprScore(
  DATA,
  n_genes = 25,
  method='UCell'
)

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='hMEs', # plot the hMEs
  order=TRUE # order so the points with highest hMEs are on top
)

# stitch together with patchwork
pdf("hMEs_OLs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='scores', # plot the hub gene scores
  order='shuffle', # order so cells are shuffled
  ucell = TRUE # depending on Seurat vs UCell for gene scoring
)
pdf("Hub_OLs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

# stitch together with patchwork
patchwork::wrap_plots(plot_list, ncol=6)

saveRDS(DATA, "OLs_wgcna.rds")
#DATA <- readRDS("Microglia_wgcna.rds")

Ctrl <- DATA@meta.data %>% subset(Celltypes == 'OLs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'OLs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

head(DMEs)

pdf("DME_lolli_OLs.pdf", width =10, height = 7)
PlotDMEsLollipop(
  DATA, 
  DMEs, 
  wgcna_name='AD_treatment', 
  pvalue = "p_val_adj")
dev.off()

pdf("DME_vol_OLs.pdf", width =10, height = 7)
PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)
dev.off()
```

### Enrichment

```{r}
DATA <- RunEnrichr(
  DATA,
  dbs=c("GO_Biological_Process_2025", "Reactome_Pathways_2024"),
  max_genes = Inf # use max_genes = Inf to choose all genes
)

# retrieve the output table
enrich_df <- GetEnrichrTable(DATA)

# look at the results
head(enrich_df)
write.csv(enrich_df, "OL_enrich.csv")
pdf("OL_enrich.pdf")
EnrichrBarPlot(
  DATA,
  outdir = "enrichr_plots", # name of output directory
  n_terms = 10, # number of enriched terms to show (sometimes more are shown if there are ties)
  plot_size = c(5,7), # width, height of the output .pdfs
  logscale=TRUE # do you want to show the enrichment as a log scale?
)
dev.off()
```

## OPCs

```{r}
DATA <- readRDS("DATA_final.rds") 
DATA <- SetIdent(DATA, value = "Celltypes") 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs","MG")) 
DATA$Celltypes[WhichCells(DATA, idents = "MG")] <- "Microglia"
DATA@reductions$UMAP <- DATA@reductions$umap.harmony30
DATA@reductions$umap.harmony30 <- NULL 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs"))

DATA <- SetupForWGCNA(
  DATA,
  gene_select = "fraction", # the gene selection approach
  fraction = 0.05, # fraction of cells that a gene needs to be expressed in order to be included
  wgcna_name = "AD_treatment" # the name of the hdWGCNA experiment
)

DATA <- MetacellsByGroups(DATA, group.by=c("Celltypes", "SampleID"), reduction = "harmony",k = 25, max_shared = 10, ident.group = "Celltypes")

table(DATA$metacell_grouping)
DATA <- NormalizeMetacells(DATA)
DATA <- SetDatExpr(DATA,  group_name= "OPCs",group.by = "Celltypes",assay = "RNA", layer = 'data')
DATA <- TestSoftPowers(DATA, networkType = 'signed')
plot_list <- PlotSoftPowers(DATA)

patchwork::wrap_plots(plot_list, ncol=2)

DATA <- ConstructNetwork(
  DATA,
  tom_name = 'OPCs_tom', # name of the topoligical overlap matrix written to disk
  overwrite_tom = T
) #soft power =7

# compute all MEs in the full single-cell dataset
DATA <- ModuleEigengenes(
 DATA,
 group.by.vars="SampleID"
)

DATA <- ModuleConnectivity(DATA, group.by = "Celltypes",group_name = 'OPCs')
DATA <- ResetModuleNames(
  DATA,
  new_name = "OLs-M"
)
PlotKMEs(DATA, ncol=5)

library(UCell)
DATA <- ModuleExprScore(
  DATA,
  n_genes = 25,
  method='UCell'
)

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='hMEs', # plot the hMEs
  order=TRUE # order so the points with highest hMEs are on top
)

# stitch together with patchwork
pdf("hMEs_OPCs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

plot_list <- ModuleFeaturePlot(
  DATA,reduction = "UMAP",
  features='scores', # plot the hub gene scores
  order='shuffle', # order so cells are shuffled
  ucell = TRUE # depending on Seurat vs UCell for gene scoring
)
pdf("Hub_OPCs.pdf", width = 10, height = 6)
patchwork::wrap_plots(c(plot_list,DimPlot(DATA,reduction = "UMAP", group.by = "Celltypes")), ncol=3)
dev.off()

# stitch together with patchwork
patchwork::wrap_plots(plot_list, ncol=6)

saveRDS(DATA, "OPCs_wgcna.rds")
#DATA <- readRDS("Microglia_wgcna.rds")

Ctrl <- DATA@meta.data %>% subset(Celltypes == 'OPCs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'OPCs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

head(DMEs)

pdf("DME_lolli_OPCs.pdf", width =10, height = 7)
PlotDMEsLollipop(
  DATA, 
  DMEs, 
  wgcna_name='AD_treatment', 
  pvalue = "p_val_adj")
dev.off()

pdf("DME_vol_OPCs.pdf", width =10, height = 7)
PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)
dev.off()
```

### Enrichment

```{r}
DATA <- RunEnrichr(
  DATA,
  dbs=c("GO_Biological_Process_2025", "Reactome_Pathways_2024"),
  max_genes = Inf # use max_genes = Inf to choose all genes
)

# retrieve the output table
enrich_df <- GetEnrichrTable(DATA)

# look at the results
head(enrich_df)
write.csv(enrich_df, "OPCs_enrich.csv")
pdf("OPCs_enrich.pdf")
EnrichrBarPlot(
  DATA,
  outdir = "enrichr_plots", # name of output directory
  n_terms = 10, # number of enriched terms to show (sometimes more are shown if there are ties)
  plot_size = c(5,7), # width, height of the output .pdfs
  logscale=TRUE # do you want to show the enrichment as a log scale?
)
dev.off()
```

## geneNMF

```{r IbNs}
library("GeneNMF")

DATA <- readRDS("DATA_final.rds") 
DATA <- SetIdent(DATA, value = "Celltypes") 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs","MG")) 
DATA$Celltypes[WhichCells(DATA, idents = "MG")] <- "Microglia"
DATA@reductions$UMAP <- DATA@reductions$umap.harmony30
DATA@reductions$umap.harmony30 <- NULL 
DATA$Celltypes <-factor(x= DATA$Celltypes, levels = c("INPs", "IbNs", "ExNs", "OPCs", "OLs", "ASCs", "Microglia", "BAMs", "Monocytes","PCs","ABCs", "VLMCs"))

DATA <- subset(DATA, idents = "IbNs")

# 1. Split your object by condition
data_list <- SplitObject(DATA, split.by = "Condition")

geneNMF.programs <- multiNMF(data_list, assay="RNA", k=4:10, min.exp = 0.05, )

geneNMF.metaprograms <- getMetaPrograms(geneNMF.programs,
                                        metric = "cosine",
                                        specificity.weight = 5,
                                        weight.explained = 0.5,
                                        nMP=10,
                                        min.confidence = 0.7)

pdf("IbNs_NMF.pdf", width = 10, height = 10)
plotMetaPrograms(geneNMF.metaprograms,similarity.cutoff = c(0.1,1))
dev.off()

DATA <- AddModuleScore_UCell(
  obj = DATA, 
  features = geneNMF.metaprograms$metaprograms.genes$MP9,
  name = "_MP9" # This adds a suffix to the metadata column names
)

DATA <- AddModuleScore_UCell(
  obj = DATA, 
  features = geneNMF.metaprograms$metaprograms.genes$MP10,
  name = "_MP10" # This adds a suffix to the metadata column names
)

DATA <- AddModuleScore_UCell(
  obj = DATA, 
  features = geneNMF.metaprograms$metaprograms.genes$MP6,
  name = "_MP6" # This adds a suffix to the metadata column names
)


pdf("IbNs_NMF_score.pdf", width = 5, height = 3)
VlnPlot(DATA, features = c("signature_1_MP6","signature_1_MP9"), group.by = "Condition")
VlnPlot(DATA, features = c("signature_1_MP6","signature_1_MP9"), group.by = "Condition", pt.size = 0)
DotPlot(DATA, features = c("signature_1_MP6","signature_1_MP9"), group.by = "Condition") + coord_flip()
dev.off()

pdf("IbNs_MP6.pdf", width = 8, height = 9)
VlnPlot(DATA, features = c("Cpa6", "Dnah11", "Adamts19", "Lgr6", "Lgr5", "Vipr2", "Syt10", "ENSMUSG00000120491", "Kank3", "Gcnt1", "Stard8", "Gm20658", "Pparg", "Ccn4", "Frmd7", "Trdn", "2900079G21Rik", "9330158H04Rik", "Meis1", "Crhr1", "Ptpn14", "Shisa8", "A930009A15Rik", "Rai14", "Tshz1", "Cpne4", "Csgalnact1", "Abi3bp", "Fbn2", "Plekha2", "Pbx3"), group.by = "Condition", stack = T, flip = T)
dev.off()

pdf("IbNs_MP6_2.pdf", width = 12, height = 9)
VlnPlot(DATA, features = c("Cpa6", "Dnah11", "Adamts19", "Lgr6", "Lgr5", "Vipr2", "Syt10", "ENSMUSG00000120491", "Kank3", "Gcnt1", "Stard8", "Gm20658", "Pparg", "Ccn4", "Frmd7", "Trdn", "2900079G21Rik", "9330158H04Rik", "Meis1", "Crhr1", "Ptpn14", "Shisa8", "A930009A15Rik", "Rai14", "Tshz1", "Cpne4", "Csgalnact1", "Abi3bp", "Fbn2", "Plekha2", "Pbx3"), group.by = "SampleID", stack = T, flip = T)
dev.off()
```

## Aggregate

```{r}

DATA <- readRDS("ExNs_wgcna.rds")
Ctrl <- DATA@meta.data %>% subset(Celltypes == 'ExNs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'ExNs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

ExNs <- PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)

ExNs <- ExNs@data %>% mutate(Celltypes = "ExNs")

DATA <- readRDS("IbNs_wgcna.rds")
Ctrl <- DATA@meta.data %>% subset(Celltypes == 'IbNs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'IbNs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

IbNs <- PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)

IbNs <- IbNs@data %>% mutate(Celltypes = "IbNs")

DATA <- readRDS("ASCs_wgcna.rds")
Ctrl <- DATA@meta.data %>% subset(Celltypes == 'ASCs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'ASCs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

ASCs <- PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)

ASCs <- ASCs@data %>% mutate(Celltypes = "ASCs")


DATA <- readRDS("OLs_wgcna.rds")
Ctrl <- DATA@meta.data %>% subset(Celltypes == 'OLs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'OLs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

OLs <- PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)

OLs <- OLs@data %>% mutate(Celltypes = "OLs")

DATA <- readRDS("Microglia_wgcna.rds")
Ctrl <- DATA@meta.data %>% subset(Celltypes == 'Microglia' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'Microglia' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

Microglia <- PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)

Microglia <- Microglia@data %>% mutate(Celltypes = "Microglia")

DATA <- readRDS("OPCs_wgcna.rds")
Ctrl <- DATA@meta.data %>% subset(Celltypes == 'OPCs' & Condition == 'Mu93') %>% rownames
Treated <- DATA@meta.data %>% subset(Celltypes == 'OPCs' & Condition == 'MuLI') %>% rownames

DMEs <- FindDMEs(
  DATA,
  barcodes1 = Ctrl,
  barcodes2 = Treated,
  test.use='wilcox',
  wgcna_name='AD_treatment'
)

OPCs <- PlotDMEsVolcano(
  DATA,
  DMEs,
  wgcna_name = 'AD_treatment'
)

OPCs <- OPCs@data %>% mutate(Celltypes = "OPCs")

```

```{r}
data <- bind_rows(ExNs, IbNs, Microglia, ASCs, OLs, OPCs)
data <- data %>%
  mutate(
    neglog10_p = -log10(p_val_adj),
    sig = neglog10_p >= 1.3 & abs(avg_log2FC) >= 0.5
  )



ggplot(data, aes(x = avg_log2FC, y = -log10(p_val_adj))) +

  # 🔹 Grey background for non-significant region
  annotate(
    "rect",
    xmin = -Inf, xmax = Inf,
    ymin = -Inf, ymax = 1.3,
    fill = "grey80",
    alpha = 0.5
  ) +

  # 🔹 Points
   geom_point(
    data = data %>% filter(!sig),
    aes(color = Celltypes),
    size = 1.5,
    alpha = 0.6
  ) +

  # 🔹 highlighted points (significant)
  geom_point(
    data = data %>% filter(sig),
    aes(color = Celltypes),
    size = 3,
    alpha = 0.9
  ) +
  ggrepel::geom_text_repel(
    data = data%>% filter(sig),
    aes(label = anno),
    size = 3
  )+

  # 🔹 Vertical dashed lines (logFC thresholds)
  geom_vline(xintercept = c(-0.5, 0.5),
             linetype = "dashed",
             color = "grey") +

  # 🔹 Horizontal dashed line (p-value threshold)
  #geom_hline(yintercept = 1.3,linetype = "dashed",color = "grey") +

  # 🔹 Style
  theme_classic() +
  labs(
    x = "avg_log2FC",
    y = "-log10(adj. p-value)",
    title = "Combined Volcano Plot"
  )

ggsave("combined_volcano.pdf", width = 8, height = 6)
```


